---
title: "Extractive Industry Trends"
author: "Matthew Ross"
date: "November 18, 2016"
output: 
  html_document:
    toc: true
    toc_depth: 3
---

This document serves as documentation of exploratory data analysis for the extractive industries portion of the Intergovernmental Panel on Biodiversity and Ecosystem Services global report on land-use change. For purposes of this report we bin extractive industries into two categories

- Fuels (including: coal, natural gas, oil, and uranium/plutonium mining)

- Mineral resources(including: rare earth minerals, metals, iron, etc...)

First let's look at global trends in coal

#Fuels

##Coal

```{r libraries,include=F}
library(tidyverse)
```

This data comes from the Energy Information Administration [website](https://www.eia.gov/beta/international/data/browser/#/?pa=0000000000000000000000000000000000000000000000000000g&c=ruvvvvvfvtvnvv1urvvvvfvvvvvvfvvvou20evvvvvvvvvnvvuvo&ct=0&tl_id=1-A&vs=INTL.7-1-AFG-TST.A&ord=SA&vo=0&v=H&start=1980&end=2014)
```{r}
#Read in data and Clean it up a little. Units are in imperial short tons. Need to convert to metric tons. 
#Also switching format to be in tidy data format (long). #Remove countries that produce no coal
coal <- read.csv('data/totalcoal.csv',stringsAsFactors = F) %>%
  gather(key='Year',value='Coal.kst',-country) %>%
  mutate(Year=as.numeric(gsub('X','',.$Year))) %>%
  mutate(Coal.kst=as.numeric(Coal.kst)) %>%
  filter(!is.na(Coal.kst))

globeplot <- coal %>% filter(country=='GlobalTotal') %>%
  ggplot(aes(x=Year,y=Coal.kst)) + geom_point() 
globeplot

```

